---
title: "July 4th Analysis"
author: "Dillon"
date: "6/28/2022"
output: 
  html_document:
    theme: readable
    highlight: zenburn
    toc: true
---

### Load in my go-to packages 
```{r loading, warning=F, message=F}
library(tidyverse)
library(rio)
library(janitor)
library(here)
library(lubridate)
library(mapview)
library(sf)
library(RColorBrewer)
library(DT)
```

### Load in the data we need for the analysis 
```{r results=F, warning=F, message=F}

daily <- read_csv(here("data", "daily_avg_by_device.csv")) %>% 
  clean_names()

hourly <- read_csv(here("data", "hourly_avg_complete.csv")) %>% 
  clean_names()

friendly_names <-  
  read_csv("friendly_names_and_ids.csv") %>% 
  clean_names() %>% 
  select(msr_device_nbr, community, device_friendly_name)


```
### Join devices to device street names and community area names 
```{r echo=F}
#Join the hourly numbers with device street names and community area names so that we can see where these lats and longs actually are
names_hourly <- 
  hourly %>% 
  inner_join(friendly_names, by = "msr_device_nbr")

```
### Make a CSV for Charmaine to work on plots 
```{r echo=F}
#Make the csv that will include summer hourly averages for every hour of the day compared with hourly averages on July 4th and July 5th morning 
summer_hourly_average <- 
  hourly %>% 
  filter(between(reading_date_local, as.Date("2021-07-21"), as.Date("2021-10-31"))) %>% 
  group_by(reading_hour_local) %>% 
  summarize(summer_avg = mean(avg_calibrated_pm25))

line_chart_df <- 
  hourly %>%
  filter(reading_date_local %in% as.Date(c("2021-07-04","2021-07-03", "2021-07-05"))) %>% 
  group_by(reading_date_local, reading_hour_local) %>% 
  summarize(avg = mean(avg_calibrated_pm25)) %>% 
  pivot_wider(names_from = reading_date_local, values_from = avg) %>% 
  inner_join(summer_hourly_average, by = "reading_hour_local")

export(line_chart_df, "line_chart.csv")

```

### CQ: "At about 8 p.m. on the Fourth of July last year, air pollution levels across Chicago started to climb — rapidly. Just a few hours later, as Chicagoans watched brightly-colored fireworks explode in the sky, the city reached a level of air pollution more than five times the hourly average of a normal summer evening."


```{r echo=F}
evening_hourly_average <- 
  hourly %>% 
  filter(reading_date_local == as.Date("2021-07-04")) %>% 
  filter(reading_hour_local %in% c("20","21","22","23","0", "1")) %>% 
  group_by(reading_date_local, reading_hour_local) %>% 
  summarize(avg = mean(avg_calibrated_pm25)) %>% 
  inner_join(summer_hourly_average, by = "reading_hour_local")

datatable(evening_hourly_average, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))

```

### CQ: "The hours between 9 p.m. and midnight on the Fourth of July 2021 ranked as the most intensely polluted hours the city experienced at any point over the past year." 

Note: December 16, 2021 was a very windy winter morning and the "pollution" only last a single hour, a data point that Madeline Daepp wasn't familiar with, so we can consider that a not worth digging into right now. The other most polluted hours are the wildfire days in late July 
```{r echo=F}

hourly_comparison <- 
  hourly %>% 
  group_by(reading_date_local, reading_hour_local) %>% 
  summarize(mean = mean(avg_calibrated_pm25)) %>% 
  filter(mean > 31)

datatable(hourly_comparison, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))

```
### CQ: "During the most intense period of pollution around 10 p.m., five sensors in the network recorded hourly averages over 100 micrograms per cubic meter, according to our analysis. All of these sensors are on the city’s south side - in the Englewood, South Chicago, Washington Park, and Ashburn community areas."

```{r echo=F }
highest_sensors_10_pm <-  
  names_hourly %>% 
  filter(reading_date_local == as.Date("2021-07-04")) %>% 
  filter(reading_hour_local %in% c("22")) %>% 
  filter(avg_calibrated_pm25 > 100)

datatable(highest_sensors_10_pm, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))
  
```

```{r}

mapview(highest_sensors_10_pm, xcol = "longitude", ycol = "latitude", zcol = "avg_calibrated_pm25", col.regions=brewer.pal(10, "YlOrRd"), crs = 4269, grid = FALSE)

```

  
### CQ: "The only area in the city that experienced more intense pollution was in Austin around 9 p.m., where a sensor at the Harrison and Central bus stop right next to Columbus Park reached an hourly average 149 micrograms per cubic meter."


```{r echo=F}
highest_sensors_all <-  
  hourly %>% 
  filter(reading_date_local == as.Date("2021-07-04")) %>% 
  filter(avg_calibrated_pm25 > 100)

datatable(highest_sensors_all, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))


```

```{r}
mapview(highest_sensors_all, xcol = "longitude", ycol = "latitude", zcol = "avg_calibrated_pm25", col.regions=brewer.pal(10, "YlOrRd"), crs = 4269, grid = FALSE)


```
